---
title: "Airlines"
author: "Alena Zhukouskaya"
date: "26 04 2022"
output: pdf_document
---

\begin{center}
Uni Databases Project - Airlines
\end{center}
\begin{center}Project presenting commands together with generated tables from the analysis of the database on flight delays in the USA in July 2017
\end{center}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("odbc")
install.packages("DBI")
install.packages("keyring")
library(odbc)
library(DBI)
library(keyring)
```

```{r, echo=FALSE}
con <- dbConnect(drv = odbc(),
                 Driver   = "SQL Server",
                 Server   = "Server",
                 Database = "Database",
                 UID      = "UID",
                 PWD      = key_get("", ""),
                 Port     = Port)
```

### Question 1: What was the average delay?

```{sql connection=con}
  select avg(FD.arr_delay_new) as avg_delay
  from Flight_delays FD
```

### Question 2: What was the maximum delay?

```{sql connection=con}
  select max(FD.arr_delay_new) as max_delay
  from Flight_delays FD
```

### Question 3: Which flight was delayed the most?

```{sql connection=con}
  select FD.fl_date, FD.arr_delay_new, FD.carrier,  FD.origin_city_name, FD.dest_city_name
    from Flight_delays FD
    where FD.arr_delay_new=(select max(arr_delay_new)
    from Flight_delays)
```

### Question 4: Which days of the week are the worst to travel?

```{sql connection=con}
  select 
    case FD.day_of_week
    when 1 then 'Monday'
    when 2 then 'Tuesday'
    when 3 then 'Wednesday'
    when 4 then 'Thursday'
    when 5 then 'Friday'
    when 6 then 'Saturday'
    when 7 then 'Sunday'
    end as weekday_name,
    avg(FD.arr_delay_new) as avg_delay
  from Flight_delays FD
  group by FD.day_of_week
  order by avg(FD.arr_delay_new) desc
```

### Question 5: Which airlines flying from San Francisco (SFO) have the fewest arrival delays?

```{sql connection=con}
  select avg(all isnull(FD.arr_delay_new,0)) as 'avg_delay', A.airline_name
  from Flight_delays FD inner join Airlines A 
    on FD.airline_id=A.airline_id
  where A.airline_id in (
    select airline_id
    from Flight_delays
    where origin='SFO')
  group by A.airline_name
  order by avg(all isnull(FD.arr_delay_new,0)) desc
```

### Question 6: What part of the airline has regular delays, i.e. its flight has an average of at least 10 minutes delays?

```{sql connection=con}
  select cast(count(FD.ile) as real)/(select count(distinct airline_id)
    from Flight_delays) as late_proportion
  from (select count(*) as 'how many'
    from Flight_delays
    group by airline_id
    having avg(arr_delay_new)>9) as FD

```

### Question 7: How do departure delays affect arrival delays?

```{sql connection=con}
  SELECT (Avg(dep_delay_new*arr_delay_new)-(Avg(dep_delay_new)*Avg(arr_delay_new)))/
  (StDevP(dep_delay_new)*StDevP(arr_delay_new))
  as [Pearsons r]
  from Flight_delays
```

### Question 8: Which airline had the biggest increase (difference) in average delay of arrivals in the last week of the month, i.e. between 1-23 and 24-31 July?

```{sql connection=con}
  select top 1 dt.sr-pt.sr as [delay_increase], A.airline_name
  from (select airline_id, avg(arr_delay_new) as sr
    from Flight_delays
    where day_of_month <= 23 and month=7
    group by airline_id) as pt
  inner join
    (select airline_id, avg(arr_delay_new) as sr
    from Flight_delays
    where day_of_month >=24 and day_of_month <=31 and month=7
    group by airline_id) as dt
  on pt.airline_id=dt.airline_id
  inner join Airlines A
  on dt.airline_id=A.airline_id
  order by dt.sr-pt.sr desc
```

### Question 9: Which airlines fly both SFO → PDX (Portland) and SFO → Eugene (Eugene)?

```{sql connection=con}
  select distinct A.airline_name
  from Flight_delays FD inner join Airlines A
    on FD.airline_id=A.airline_id
  where (FD.origin='SFO' and FD.dest='PDX') and exists (
    select airline_id
    from Flight_delays
    where airline_id=FD.airline_id and origin='SFO' and dest='EUG')
```

### Question 10: What is the fastest way to get from Chicago to Stanford, assuming departure after 14:00 local time?

```{sql connection=con}
  select avg(FD.arr_delay_new) as 'avg_delay', FD.origin, FD.dest
  from Flight_delays FD
  where (FD.origin='MDW' or FD.origin='ORD')
    and FD.crs_dep_time>1400
    and (FD.dest='SFO' or FD.dest='SJC' or FD.dest='OAK')
  group by FD.origin, FD.dest
  order by avg(FD.arr_delay_new) desc
```